Particle rejuvenation of Rao-Blackwellized Sequential Monte Carlo smoothers for Conditionally Linear and Gaussian models

### 2017-07-03 TexSubmission/clgm.tex
Submitted version

\include{clgm_def}
\begin{document}

@@

\subsection{Numerical results}
The initial transition probability in CMA-ES is chosen as $Q(1,1) = 0.98, \; Q(2,2) = 0.97$ and $\pi_1 = \pi_2 = 0.5$ where the number $1$ represents the backwardation regime and $2$ represents the contango regime. The other parameters are initialized as shown in the following table:
\begin{figure}[h!]
	\centering
	\begin{tabular}{||c|c|c|c|c|c|c|c|c|c|c|c|c||}
		$\kappa$ & $\alpha_1$ & $\alpha_2$ & $\sigma_1$ & $\sigma_2$ & $\eta_1$ & $\eta_2$ & $\rho_1$ & $\rho_2$ & $g_1$ & $g_2$ & $g_3$ & $g_4$ \\
		\hline
		5.0 & 0.1 & -0.05 & 0.4 & 0.4 & 0.5 & 0.5 & 0.75 & 0.65 & 0.1 & 0.1 & 0.1 & 0.1
	\end{tabular}
	\caption{Initial values for the EM algorithm.}
\end{figure}\\
The number of particles is set to $N = 100$, $\tau  = 1/52$. 
The interest rate is $r = 0.0296$ as in \cite{almansour:2016}. The initial guess for the mean and variance of the initial state are
\[
\mu_1 = \begin{pmatrix} \ln F^{(market)}_{1,4} & r-\cfrac{\ln F^{(market)}_{1,16}-\ln F^{(market)}_{1,4}}{(16-4)\tau} \end{pmatrix}\quad \mbox{and}\quad \Sigma_1 = \begin{pmatrix} 0.05 & 0 \\ 0 & 0.05 \end{pmatrix}\eqsp.
\]
The CMA-ES algorithm is used with an initial standard deviation for the parameters $\sigma_{cmaes} = 0.005$, a number of selected search points $\mu_{cmaes} = 20$ and a population size $\lambda_{cmaes} = 100$. The algorithm is stopped after $10000$ iterations.

@@

@@

\bibliography{./clgm_bib}
\end{document}

>>>>

\documentclass[nolayout]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath,amsthm}
\usepackage{geometry}
\usepackage[textwidth=4cm,textsize=footnotesize]{todonotes}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{amssymb,color,bbm,xargs}
\usepackage{graphicx}
\usepackage[active]{srcltx}
\usepackage{ifthen}
\usepackage{enumerate}
\usepackage{color}
\usepackage{accents}
\usepackage{dsfont}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{subfig}
\usepackage{multirow}
\usepackage{placeins}

\usepackage{aliascnt}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{assumption}{H\hspace{-3pt}}
\newtheorem*{assumption*}{A}
\newtheorem{assumptionDelta}{H\hspace{-3pt}(\Delta)}
\newtheorem{assumptionsmallSetD}{H\hspace{-3pt}(\smallSetD \times \smallSetD)}
\newtheorem{assumptionAR}{AR\hspace{-3pt}}
\newtheorem{assumptionCN}{CN\hspace{-3pt}}

\newaliascnt{proposition}{theorem}
\newtheorem{proposition}[proposition]{Proposition}
\aliascntresetthe{proposition}

\newaliascnt{lemma}{theorem}
\newtheorem{lemma}[lemma]{Lemma}
\aliascntresetthe{lemma}
\newaliascnt{corollary}{theorem}
\newtheorem{corollary}[corollary]{Corollary}
\aliascntresetthe{corollary}

\newtheorem{hypothese}{Hypothesis}
\newtheorem{fact}[theorem]{Fact}

\theoremstyle{definition}
\newaliascnt{definition}{theorem}
\newtheorem{definition}[definition]{Definition}
\aliascntresetthe{definition}

\newaliascnt{remark}{theorem}
\newtheorem{remark}[remark] {Remark}
\aliascntresetthe{remark}

\usepackage{hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}

\providecommand*{\definitionautorefname}{Definition}
\providecommand*{\lemmaautorefname}{Lemma}
\providecommand*{\exampleautorefname}{Example}
\providecommand*{\propositionautorefname}{Proposition}
\providecommand*{\procautorefname}{Procedure}
\providecommand*{\exerciseautorefname}{Exercise}
\providecommand*{\corollaryautorefname}{Corollary}
\providecommand*{\remarkautorefname}{Remark}
\providecommand*{\algorithmautorefname}{Algorithm}
\providecommand*{\assumptionautorefname}{H\hspace*{-3pt}}  
\providecommand*{\assumptionARautorefname}{AR\hspace*{-3pt}}
\providecommand*{\assumptionCNautorefname}{CN\hspace*{-3pt}}

\newcommand*{\largedot}{\cdot}

\def\Rset{\mathbb{R}}
\def\Nset{\mathbb{N}}
\def\nset{\mathbb{N}}
\def\rset{\mathbb{R}}
\def\varb{b}
\newcommand{\1}{\mathbbm{1}}

\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\ceil}[1]{\left\lceil #1 \right\rceil}

\def\ie{\textit{i.e.}}
\newcommand{\coint}[1]{\left[#1\right)}
\newcommand{\ocint}[1]{\left(#1\right]}
\newcommand{\ooint}[1]{\left(#1\right)}
\newcommand{\ccint}[1]{\left[#1\right]}
\newcommand{\Wienerspace}{\mathbf{W}}

\def\PE{\mathbb{E}}
\newcommandx{\PEt}[2][1=]{\mathbb{E}_{#1}\left[#2\right]}
\def\rmd{\mathrm{d}}
\def\rme{\mathrm{e}}
\def\target{\pi}
\def\dim{d}
\def\score{\accentset{\largedot}{\ell}}
\def\Vdot{\dot{V}}
\def\unitarget{f}
\def\rmL{\mathrm{L}}
\def\dimy{\mathsf{p}}
\def\dimz{\mathsf{m}}
\def\bfX{\boldsymbol{X}}
\def\xbf{\boldsymbol{x}}
\def\ybf{\boldsymbol{y}}
\def\Xbf{\boldsymbol{X}}
\def\G{\mathcal{G}}
\def\barG{\overline{G}}
\def\barH{\overline{H}}
\def\Gammabf{\mathbf{\Gamma}}
\def\Ybf{\boldsymbol{Y}}
\newcommand{\norm}[1]{\left\Vert #1 \right \Vert}
\newcommand{\plusinfty}{+ \infty}
\newcommand{\defEns}[1]{\left \{ #1 \right \}}
\newcommand{\abs}[1]{\left\vert #1 \right\vert}
\def\Zbf{\boldsymbol{Z}}
\def\eqsp{\,}
\def\leb{\lambda}
\def\limd{\underset{d\to +\infty}{\longrightarrow}}
\newcommand{\normMat}[2]{\left\|#2\right\|_{#1}}

\newcommand{\parenthese}[1]{\left( #1 \right)}
\newcounter{hypH}
\newenvironment{hypH}{\refstepcounter{hypH}\begin{itemize}
\item[{\bf H\arabic{hypH}}]}{\end{itemize}}
\newcommand{\eqdef}{\ensuremath{:=}}
\newcommand{\setAccept}{\mathcal{A}}
\def\iid{\operatorname{i.i.d.}}
\newcommand{\sachant}[1]{\left| #1 \right.}
\newcommand{\setDisconDotV}{\mathcal{D}_{\dot{V}}}

\def\restriction#1#2{\mathchoice
              {\setbox1\hbox{${\displaystyle #1}_{\scriptstyle #2}$}
              \restrictionaux{#1}{#2}}
              {\setbox1\hbox{${\textstyle #1}_{\scriptstyle #2}$}
              \restrictionaux{#1}{#2}}
              {\setbox1\hbox{${\scriptstyle #1}_{\scriptscriptstyle #2}$}
              \restrictionaux{#1}{#2}}
              {\setbox1\hbox{${\scriptscriptstyle #1}_{\scriptscriptstyle #2}$}
              \restrictionaux{#1}{#2}}}
\def\restrictionaux#1#2{{#1\,\smash{\vrule height .8\ht1 depth .85\dp1}}_{\,#2}}

\newcommandx\sequence[3][2=,3=]
{\ifthenelse{\equal{#3}{}}{\ensuremath{\{ #1_{#2}\}}}{\ensuremath{( #1_{#2}, \eqsp #2 \in #3 )}}}
\newcommandx{\sequencen}[2][2=n\in\nset]{\ensuremath{( #1, \eqsp #2 )}}
\newcommandx\dsequence[4][3=k,4=\zset]{\ensuremath{( (#1_{#3}, #2_{#3}), \eqsp #3 \in #4 )}}
\newcommandx{\as}[1][1=\PP]{\ensuremath{#1\--\mathrm{a.s.}}}
\newcommand{\eqspp}{\ \ }

\def\canonicalFiltration{\mathscr{B}}

\def\tildeX{\widetilde{X}}
\def\tildeY{\widetilde{Y}}
\def\barX{\bar{X}}
\def\barY{\bar{Y}}
\def\barS{\bar{S}}
\newcommand{\coupling}[2]{\mathcal{C}(#1,#2)}
\newcommand\proba[1]{\mathbb{P}\left[ #1 \right]}
\newcommand\expe[1]{\mathbb{E}\left[ #1 \right]}
\def\osc{\operatorname{osc}}
\def\borel{\mathcal{B}}
\def\mcf{\mathcal{F}}

\begin{document}

@@

\subsection{Numerical results}
The initial transition probability in CMA-ES is chosen as $Q(1,1) = 0.98, \; Q(2,2) = 0.97$ and $\pi_1 = \pi_2 = 0.5$ where the number $1$ represents the backwardation regime and $2$ represents the contango regime. The other parameters are initialized as shown in Table~\ref{fig:CMEinitialestimates}.
\begin{table}[h!]
	\centering
	\begin{tabular}{||c|c|c|c|c|c|c|c|c|c|c|c|c||}
		$\kappa$ & $\alpha_1$ & $\alpha_2$ & $\sigma_1$ & $\sigma_2$ & $\eta_1$ & $\eta_2$ & $\rho_1$ & $\rho_2$ & $g_1$ & $g_2$ & $g_3$ & $g_4$ \\
		\hline
		5.0 & 0.1 & -0.05 & 0.4 & 0.4 & 0.5 & 0.5 & 0.75 & 0.65 & 0.1 & 0.1 & 0.1 & 0.1
	\end{tabular}
	\caption{Initial values for the EM algorithm.}
	\label{fig:CMEinitialestimates}
\end{table}
The number of particles is set to $N = 100$, $\tau  = 1/52$. 
The interest rate is $r = 0.0296$ as in \cite{almansour:2016}. The initial guess for the mean and variance of the initial state are
\[
\mu_1 = \begin{pmatrix} \ln F^{(market)}_{1,4} & r-\cfrac{\ln F^{(market)}_{1,16}-\ln F^{(market)}_{1,4}}{(16-4)\tau} \end{pmatrix}\quad \mbox{and}\quad \Sigma_1 = \begin{pmatrix} 0.05 & 0 \\ 0 & 0.05 \end{pmatrix}\eqsp.
\]
The CMA-ES algorithm is used with an initial standard deviation for the parameters $\sigma_{cmaes} = 0.005$, a number of selected search points $\mu_{cmaes} = 20$ and a population size $\lambda_{cmaes} = 100$. The algorithm is stopped after $10000$ iterations.

@@

\subsection*{Competing interests}
The authors have no competing interests.

\subsection*{Funding}
This work has been developed during a three year Ph.D. at T\'el\'ecom ParisTech and Lunalogic.

\subsection*{Authors' contributions}
All the authors have contributed to the conception of the algorithms and to the redaction of the manuscript.

@@

\newcommand{\etalchar}[1]{$^{#1}$}
\begin{thebibliography}{FGDW02}

\bibitem[Alm16]{almansour:2016}
A.~Almansour.
\newblock Convenience yield in commodity price modeling: A regime switching
  approach.
\newblock {\em Energy Economics}, 53:238--247, 2016.

\bibitem[Bar06]{barber:2006}
D.~Barber.
\newblock {Expectation correction for smoothed inference in switching linear
  dynamical systems}.
\newblock {\em Journal of {M}achine {L}earning {R}esearch}, 7:2515--2540, 2006.

\bibitem[BDM10]{briers:doucet:maskell:2010}
M.~Briers, A.~Doucet, and S.~Maskell.
\newblock Smoothing algorithms for state-space models.
\newblock {\em {A}nn. {I}nst. {S}tat. {M}ath}, 62(1):61--89, 2010.

\bibitem[BF63]{bryson:frazier:1963}
A.E. Bryson and M.~Frazier.
\newblock {Smoothing for linear and nonlinear dynamic systems}.
\newblock {\em {Proc. Optimum Sys. Synthesis Conf.}}, 1963.

\bibitem[BGM08]{barembruch:garivier:moulines:2008}
S.~Barembruch, A.~Garivier, and E.~Moulines.
\newblock {On optimal sampling for particle filtering in digital
  communication}.
\newblock {\em IEEE 9th Workshop on Signal Processing Advances in Wireless},
  pages 634--638, 2008.

\bibitem[CL00]{chen:liu:2000}
R.~Chen and J.S. Liu.
\newblock Mixture kalman filters.
\newblock {\em Journal of the Royal Statistical Society B}, 62:493--508, 2000.

\bibitem[DGA00]{doucet:godsill:andrieu:2000}
A.~Doucet, S.~Godsill, and C.~Andrieu.
\newblock On sequential monte carlo sampling methods for bayesian filtering.
\newblock {\em Statistics and computing}, 10:197--208, 2000.

\bibitem[DGK01]{doucet:gordon:krishnamurthy:2001}
A.~Doucet, N.~Gordon, and V.~Krishnamurthy.
\newblock Particle filters for state estimation of jump {M}arkov linear
  systems.
\newblock {\em {IEEE T}ransactions on {S}ignal {P}rocessing}, 49(3):613--624,
  2001.

\bibitem[DLR77]{dempster:laird:rubin:1977}
A.~P. Dempster, N.~M. Laird, and D.~B. Rubin.
\newblock Maximum likelihood from incomplete data via the {EM} algorithm.
\newblock {\em Journal of the Royal Statistical Society B}, 39(1):1--38 (with
  discussion), 1977.

\bibitem[DM13]{delmoral:2013}
P.~Del~Moral.
\newblock {\em Mean field simulation for Monte Carlo integration}.
\newblock Chapman \& Hall / CRC Monographs on Statistics \& Applied
  Probability, 2013.

\bibitem[FC03]{fearnhead:clifford:2003}
P.~Fearnhead and P.~Clifford.
\newblock {On-line inference for hidden {M}arkov models via particle filters}.
\newblock {\em {J.R. Statist. Soc. B}}, 65:887--899, 2003.

\bibitem[FGDW02]{fong:godsill:doucet:west:2002}
W.~Fong, S.J. Godsill, A.~Doucet, and M.~West.
\newblock {Monte Carlo smoothing with application to audio signal enhancement}.
\newblock {\em IEEE Transactions on {S}ignal {P}rocessing}, 50(2):438--449,
  2002.

\bibitem[FWT10]{fearnhead:wyncoll:tawn:2010}
P.~Fearnhead, D.~Wyncoll, and J.~Tawn.
\newblock {A sequential smoothing algorithm with linear computational cost}.
\newblock {\em {Biometrika}}, 97:447--464, 2010.

\bibitem[GE90]{gibson:schwartz:1990}
R.~Gibson and Schwartz E.S.
\newblock Stochastic convenience yield and the pricing of oil contingent
  claims.
\newblock {\em Journal of finance}, 45(3):959--976, 1990.

\bibitem[HK98]{huerzeler:kunsch:1998}
M.~H{\"u}rzeler and H.~R. K{\"u}nsch.
\newblock {M}onte {C}arlo approximations for general state-space models.
\newblock {\em J. Comput. Graph. Statist.}, 7:175--193, 1998.

\bibitem[HK04]{hansen:kern:2004}
N.~Hansen and S.~Kern.
\newblock {Evaluating the CMA Evolution Strategy on Multimodal Test Functions}.
\newblock {\em Eighth International Conference on Parallel Problem Solving from
  Nature}, 72:337--354, 2004.

\bibitem[HO01]{hansen:ostermeier:2001}
N.~Hansen and A.~Ostermeier.
\newblock {Completely derandomized self-adaptation in evolution strategies}.
\newblock {\em Evolutionary {C}omputation}, 9(2):159--195, 2001.

\bibitem[Kal60]{kalman:1960}
R.E. Kalman.
\newblock {A new approach to linear filtering and prediction problems}.
\newblock {\em J. {B}asic {E}ngng}, 82:35--45, 1960.

\bibitem[Kim94]{kim:1994}
CJ. Kim.
\newblock Dynamic linear models with markov-switching.
\newblock {\em Journal of Econometrics}, 60(1-2):1--22, 1994.

\bibitem[LBGS13]{lindsten:bunch:godsill:schon:2013}
F.~Lindsten, P.~Bunch, S.J. Godsill, and T.B. Schon.
\newblock Rao-{B}lackwellized particle smoothers for mixed linear/nonlinear
  state-space models.
\newblock {\em Proceedings of the 38th {IEEE I}nternational {C}onference on
  {A}coustics, {S}peech and {S}ignal {P}rocessing ({ICASSP})}, 2013.

\bibitem[LBS{\etalchar{+}}16]{lindsten:bunch:sarkka:schon:godsill:2015}
F.~Lindsten, P.~Bunch, S.~Sarkka, T.B. Schon, and S.J. Godsill.
\newblock Rao-{B}lackwellized particle smoothers for conditionally linear
  {G}aussian models.
\newblock {\em IEEE {J}ournal of {S}elected {T}opics in {S}ignal {P}rocessing},
  10(2):353--365, 2016.

\bibitem[RST65]{rauch:striebel:tung:1965}
A.E. Rauch, C.T. Striebel, and F.~Tung.
\newblock {Maximum likelihood estimates of linear dynamic systems}.
\newblock {\em American Institute of Aeronautics and Astronautics journal},
  3(8):1445--1450, 1965.

\bibitem[Sar13]{sarkka:2013}
S.~Sarkka.
\newblock {\em Bayesian filtering and smoothing}.
\newblock Cambridge {U}niversity {P}ress, 2013.

\bibitem[SBG12]{sarkka:bunch:godsill:2012}
S.~Sarkka, P.~Bunch, and S.J. Godsill.
\newblock A backward-simulation based {R}ao-{B}lackwellized particle smoother
  for conditionally linear {G}aussian models.
\newblock {\em Proceedings of the 16th {IFAC S}ymposium on {S}ystem
  {I}dentification {(SYSID)}}, 2012.

\bibitem[Sch97]{schwartz:1997}
E.S. Schwartz.
\newblock The stochastic behaviour of commodity prices: Implications for
  pricing and hedging.
\newblock {\em The {J}ournal of {F}inance}, 3:923--973, 1997.

\bibitem[SGN05]{schon:gustafsson:nordlund:2005}
T.~Schon, F.~Gustafsson, and P.-J. Nordlund.
\newblock Marginalized particle filters for mixed linear/nonlinear state-space
  models.
\newblock {\em {IEEE T}ransactions on {S}ignal {P}rocessing}, 53(7):2279--2289,
  2005.

\end{thebibliography}

%\bibliography{./clgm_bib}
\end{document}
